---
title: "Assignment"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidytext)
```

## Reading emails
Define a function to extract the email header, any line before the first empty line occured is considered as the email header.
```{r}
# Get the email header
getHeader <- function(lines){
  index <- 1
  for(line in lines){
    line <- trimws(line)
    if(line==""){
      break
    }
    index <- index + 1
  }
  header <- lines[1:(index-1)]
  header
}
```

Define a function to extract the email message body, any content which do not belong to the multipart is considered belong to the message body.
```{r}
# get the email body
getBody <- function(lines){
  temp <- lines[grepl("Content-Type:", lines, ignore.case = T)]
  line <- ""
  if(length(temp)>=1){
    line <- temp[1]
  }
  contentType <- str_extract(line, "(?<=Content-Type: )(.*)(?=;)")
  contentType <- unlist(strsplit(contentType, ";"))[1]
  contentType <- trimws(tolower(contentType))
  if(grepl("multipart", contentType)){
    temp <- lines[grepl("boundary=", lines, ignore.case = T)]
    temp <- tolower(temp)
    boundry = str_extract(temp[1], '(?<=\")(.*)(?=")')
    if(is.na(boundry)){
      boundry = str_extract(temp[1], 'boundary=(.*)')
      boundry <- gsub("boundary=", "", boundry, ignore.case = T)
    }
    index <- which(grepl("Content-Type: text/plain", lines, ignore.case = T))
    if(length(index)>0){
      index <- index[1]
      stop <- length(lines)
      for(i in (index+1):length(lines)){
        if(grepl(boundry, lines[i])){
          stop <- i-1
          break
        }
      }
      return(lines[(index+1):stop])
    }
    return("")
  }else{
      index <- 1
      for(line in lines){
        line <- trimws(line)
        if(line==""){
          break
        }
        index <- index + 1
      }
      body <- lines[(index+1):length(lines)]
      return(body)
  }
}
```

Loop for each email, reading the lines of each email to a list.
```{r warning=FALSE, message=F}
emails <- list()
index <- 1
paths <- c()
for(dir in list.dirs("./SpamAssassinTrain")[-1]){
  filenames <- list.files(dir)
  filenames <- paste0(dir, "/", filenames)
  paths <- c(paths, filenames)
  for(i in 1:length(filenames)){
    emails[[index]] <- readLines(filenames[i])
    index <- index + 1
  }
}
```

## Data cleaning
Find the content type of each email, most of the emails are in plain text form.
```{r}
contents <- c()
for(i in 1:length(emails)){
  lines <- emails[[i]]
  lines <- lines[grepl("Content-Type:", lines, ignore.case = T)]
  line <- ""
  if(length(lines)>=1){
    line <- lines[1]
  }
  contentType <- str_extract(line, "(?<=Content-Type: )(.*)(?=;)")
  contentType <- unlist(strsplit(contentType, ";"))[1]
  contents[i] <- trimws(tolower(contentType))
}
table(contents)
```

Whether the header states that the message is a multipart/text, i.e. with attachments.
```{r}
multipartText <- grepl("multipart", contents)
```

Get the headers of each email.
```{r}
headers <- lapply(emails, getHeader)
```

Get the message body of each email.
```{r}
bodies <- lapply(emails, getBody)
```

Whether mail is Spam (TRUE) or Ham (FALSE).
```{r}
isSpam <- !grepl("ham", paths)
table(isSpam)
```

Get the number of attachments in the message, draw a histogram of the number of lines in the body, most of the emails have a line number less than 1000.
```{r}
numLinesInBody <- sapply(bodies, length)
hist(numLinesInBody)
```

Get the number of characters in the body of the email message, the distribution is right skewed  as the histogram demonstrates.
```{r}
bodyCharacterCount <- sapply(sapply(bodies, nchar, allowNA=T), sum)
hist(bodyCharacterCount)
```

Get the subjects of the emails.
```{r}
getSubject <- function(header){
  lines <- header[grepl("Subject:", header, ignore.case = T)]
  if(length(lines)>=1){
    line = lines[1]
    line <- gsub("Subject:", "", line)
    line <- trimws(line)
    line
  }else{
    return("")
  }
}
subjects <- sapply(headers, getSubject)
```

Get the number of exclamation marks (!) in the subject of the message. Most of the emails do not have  exclamation marks.
```{r}
subjectExclamationCount <- str_count(subjects, "!")
barplot(table(subjectExclamationCount))
```

Get the number of question marks in the subject.
```{r}
subjectQuestCount <- str_count(subjects, "\\?")
barplot(table(subjectQuestCount))
```

Combine all the messages lines into a text for each message.
```{r}
body_text <- sapply(bodies, paste0, collapse=", ")
```

Get the number of dollar signs in the body of the message.
```{r}
numDollarSigns <- sapply(body_text, function(text){
  str_count(text, "\\$")
})
hist(numDollarSigns)
```

```{r}
averageWordLength <- sapply(bodies, function(lines){
  df <- tibble(line = 1:length(lines), text = lines)
  df <- df %>% unnest_tokens(word, text)
  mean(nchar(df$word))
})
```

The distribution of the average number of word length is right skewed.
```{r}
hist(averageWordLength)
```

The percentage of the characters in the body of the email that are upper case is left skewed.
```{r}
percentCapitals <- str_count(body_text, "[a-zA-Z]")/str_length(body_text)
hist(percentCapitals)
```

Whether the header of the message has an In-Reply-To field.
```{r}
isInReplyTo <- sapply(headers, function(header){
  any(grepl("In-Reply-To:", header, ignore.case = T))
})
table(isInReplyTo)
```

Whether the subject has punctuation or digits surrounded by characters, e.g. V?agra and pay1ng, but not New!
```{r}
subjectPunctuationCheck <- sapply(subjects, function(subject){
  any(grepl("[A-Za-z][[:punct:]]+[A-Za-z]", subject, ignore.case = T))
})
```

Whether the message’s header had an X-Priority or X-SmellPriority that was set to high
```{r}
isPGPsigned <- sapply(emails, function(lines){
  any(grepl("PGP-?SIGNATURE", lines, ignore.case = T)) |
    any(grepl("GPG-Key-Fingerprint", lines, ignore.case = T))
})
table(isPGPsigned)
```

Find the percentage of blanks in the subject, we can find that the distribution is right skewed.
```{r}
percentSubjectBlanks <- str_count(subjects, " ")/str_length(subjects)
hist(percentSubjectBlanks)
```

Whether the message body contains a form of the introduction Dear.
```{r}
isDear <- sapply(body_text, function(text){
  any(grepl("dear ", text, ignore.case = T))
})
table(isDear)
```

Whether the text includes a line indicating an included message as identified by the word wrote:.
```{r}
isWrote <- sapply(body_text, function(text){
  any(grepl("wrote:", text, ignore.case = T))
})
table(isWrote)
```

Whether the string Re: appears as the first word in the subject of the message.
```{r}
isRe <- grepl("RE:", subjects, ignore.case = T)
```

Combine all the data.
```{r}
data <- data.frame(
  isSpam, isRe, numLinesInBody, bodyCharacterCount, subjectExclamationCount,
  subjectQuestCount, isInReplyTo, multipartText, percentSubjectBlanks,
  isPGPsigned, isDear, subjectPunctuationCheck, 
  isWrote, averageWordLength, numDollarSigns
)
head(data)
```

## Analysis
```{r}
data %>% group_by(isSpam) %>% summarise(numLinesInBody=mean(numLinesInBody)) %>%
  ggplot(aes(x=isSpam, y=numLinesInBody, fill=isSpam)) + geom_bar(stat="identity") +
  ylab("Average number lines in body")
```

The average number of lines in a spam email is longer than that of the ham emails.

```{r}
data %>% group_by(isSpam) %>% summarise(bodyCharacterCount=mean(bodyCharacterCount, na.rm=T)) %>%
  ggplot(aes(x=isSpam, y=bodyCharacterCount, fill=isSpam)) + geom_bar(stat="identity") +
  ylab("Average number of characters in body")
```

The average number of characters in spam emails is higehr than that of the ham emails.

```{r}
data %>% group_by(isSpam) %>% summarise(numDollarSigns=mean(numDollarSigns)) %>%
  ggplot(aes(x=isSpam, y=numDollarSigns, fill=isSpam)) + geom_bar(stat="identity") +
  ylab("Average number dollar signs in body")
```

The average number dollar signs in spam emails is higher than that of the ham emails.

```{r}
data %>% group_by(isSpam) %>% summarise(percent=mean(isDear)) %>%
  ggplot(aes(x=isSpam, y=percent, fill=isSpam)) + geom_bar(stat="identity")
```

The percentage of using dear in spam emails is higher than that of the ham emails.
